---
title: "Test"
author: "DeepakKalaikadal"
date: "2022-10-30"
output: html_document
---

# Overview

This is the final project for *Practical Machine Learning* course on Coursera. This project will use data collected from fitness devices attached to waist, arm, dumb-bells of 6 participants in order to predict their exercise pattern, defined by the 'classe' variable in the available data. I have decided to use 3 models for training - **Decision Trees, Random Forest, and Gradient Boosted Tree** on the training set. The best model is used to predict a validation set partitioned from the training-data using its accuracy and out-of-sample error rate. The best model is then used to predict 20 cases defined in the Testing set.

## Importing Requisite Libraries
```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set()
#Importing requisite libraries

library(lattice)
library(ggplot2)
library(caret)
library(kernlab)
library(rattle)
library(corrplot)
set.seed(1358)
```

## Reading, Cleaning, and Partitioning  the Training Dataset

We will remove variables with little relevance to the predictions, variables with near-zero variance, and columsn with largely N/A entries. We will then partition the training set to give a validation set and a training set.

```{r}
pml_training <- read.csv("C:/Users/deepa/Downloads/pml-training.csv")
pml_testing <- read.csv("C:/Users/deepa/Downloads/pml-testing.csv")
train.raw <- pml_training
test.raw <- pml_testing

#removing columns with NA > 80%
train.clean01 <- train.raw[,colMeans(is.na(train.raw)) < .8]
#removing variables which are time-based, which are irrelevant
train.clean02 <- train.clean01[,-c(1, grep("timestamp", names(train.clean01)) )]
#removing near-zero variance variables
train.clean03 <- train.clean02[,-nearZeroVar(train.clean02)]
train.clean04 <- train.clean03[,-1]
#Partitioning the Training-Set to give a Validation-Set
partition <- createDataPartition(y=train.clean03$classe, p=0.7, list=F)
train <- train.clean04[partition,]
valid <- train.clean04[-partition,]